Hierarchical Dispatching and Overriding

### 2017-06-18 Aplas17/sections/Formalization.tex
grammar check Sec3

In this section, we present the formal model \MIM{} (\emph{\textbf{F}eatherweight \textbf{H}irarchical \textbf{J}ava}) , based on
a minimal subset of Java 8 interfaces following Featherweight Java style~\cite{Igarashi01FJ}.
The syntax, typing rules and small-step semantics are included below.

@@

Expressions can be standard constructs such as variables, method
invocation, or object creation. A novel type of expressions are path-invocations like
``$e.I::m(\overline{e})$'', meaning that the dynamically binded
implementation for method $m$ should be along the path $I$. Another
interesting expression are super-invocations, which enable a method to access an old
implementation from the specified super type. Hence a super-invocation
can only be used inside an interface definition.
Finally an expression can also be annotated with its static type. But \emph{annotated expressions are only intended for the semantic rules},
hence they are not supposed to appear in the source program. A value
``$\angl{I}\new{J}$''
is the final result of multple reduction steps evaluating an
expression. Such value represents an object instance
of $J$ with an annotated static type $I$.

For simplicity \name{} does not formalize statements like assignments and so on, because they are orthogonal features.
A program consists of a list of interface declarations, plus a single expression.

@@

The method typing rule \textsc{(T-Method)} is more interesting, since the method can either be an original implementation or hierarchical overriding.
Besides type-checking the return expression,
we further use the helper function $\mostSpecific$, again formally defined in Section~\ref{subsec:auxdefs}.
By the equation ``$\mostSpecific(m, I, J) = \{J\}$'' we define the legality of method overrides as mentioned in Section~\ref{subsec:partialoverrides}, namely if it is hierarchical overriding, it should not step over original methods, otherwise
their enclosing types would be returned instead of $J$.

Finally, \textsc{(T-Invk)} defines interface checking. The condition helps to ensure unambiguity and type soundness of the calculus. We will
introduce \mbody{} and some counter-examples later. But intuitively, if $\mbody(m,I,J)$ is defined, that means $\new{I}.J::m()$ is not
ambiguous during runtime. Therefore the condition says that if an expression type checks, it should not introduce ambiguity during runtime
in any case. The interface check is responsible to capture ambiguity during compilation.

@@

\subsection{Properties}
Previously the definitions of our model are given, now we should proceed to prove the type soundness of 
the model, which relates typing to computation. The type soundness states that, if an expression is 
well-typed, then after many reduction steps it must reduce to a value, with its annotation to be subtype of the original expression type.
Following the Featherweight Java paper~\cite{Igarashi01FJ}, the type-soundness theorem 
(Theorem~\ref{theorem_soundness}) is proved by using the standard technique of subject reduction (Theorem~\ref{theorem_subject})
and progress (Theorem~\ref{theorem_progress})~\cite{Wright1994}. In Theorem~\ref{theorem_progress} ``$\#(\overline{x})$'' denotes the number of
elements.

>>>>

In this section, we present the formal model \MIM{} (\emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}) , based on
a minimal subset of Java 8 interfaces following Featherweight Java style~\cite{Igarashi01FJ}. We use the same
conventions as FJ; $\overline{I}$ is shorthand for the (possibly empty)
list $I_1, ..., I_n$, which may be indexed by $I_i$.
The syntax, typing rules, and small-step semantics are presented below.

@@

Expressions can be standard constructs such as variables, method
invocation, or object creation. A novel type of expressions are path-invocations like
``$e.I::m(\overline{e})$'', meaning that the dynamically bound
implementation for method $m$ should be along the path $I$. Another
interesting expression are super-invocations, which enable a method to access an old
implementation from the specified super type. Hence a super-invocation
can only be used inside an interface definition.
Finally, an expression can also be annotated with its static type. But \emph{annotated expressions are only intended for the semantic rules},
hence they are not supposed to appear in the source program. A value
``$\angl{I}\new{J}$''
is the final result of multiple reduction steps evaluating an
expression. Such value represents an object instance
of $J$ with an annotated static type $I$.

For simplicity, \name{} does not formalize statements like assignments and so on because they are orthogonal features.
A program consists of a list of interface declarations, plus a single expression.

@@

The method typing rule \textsc{(T-Method)} is more interesting since the method can either be an original implementation or hierarchical overriding.
Besides type-checking the return expression,
we further use the helper function $\mostSpecific$, again formally defined in Section~\ref{subsec:auxdefs}.
By the equation ``$\mostSpecific(m, I, J) = \{J\}$'' we define the legality of method overrides as mentioned in Section~\ref{subsec:partialoverrides}, namely if it is hierarchical overriding, it should not step over original methods, otherwise
their enclosing types would be returned instead of $J$.

Finally, \textsc{(T-Invk)} defines interface checking. The condition helps to ensure unambiguity and type soundness of the calculus. We will
introduce \mbody{} and some counter-examples later. But intuitively, if $\mbody(m,I,J)$ is defined, that means $\new{I}.J::m()$ is not
ambiguous during runtime. Therefore the condition says that if an expression type checks, it should not introduce ambiguity during runtime
in any case. The interface check is responsible for capturing ambiguity during compilation.

@@

\subsection{Properties}
Previously the definitions of our model are given, now we should proceed to prove the type soundness of 
the model, which relates typing to computation. The type soundness states that, if an expression is 
well-typed, then after many reduction steps it must reduce to a value, with its annotation to be a subtype of the original expression type.
Following the Featherweight Java paper~\cite{Igarashi01FJ}, the type-soundness theorem 
(Theorem~\ref{theorem_soundness}) is proved by using the standard technique of subject reduction (Theorem~\ref{theorem_subject})
and progress (Theorem~\ref{theorem_progress})~\cite{Wright1994}. In Theorem~\ref{theorem_progress} ``$\#(\overline{x})$'' denotes the number of
elements.
